Capturing Subdiffusive Solute Dynamics and Predicting Selectivity in Nanoscale Pores with Time Series Modeling

### 2020-02-07 Ben_Manuscripts/stochastic_transport/Final_Draft.tex
HDP-HMM outline

  Next, we generalize Markov state models, treating the
  configurational state of the system as a Markov process, each
  state of which has  
  %MRS10:
  %each having separate
  distinct transport properties. For each state and transition
  between states, we parameterize the distribution and temporal correlation 
  structure of positional fluctuations as a means of characterization and to 
  allow us to predict solute MSDs. We show that both models reasonably reproduce
  the MSDs calculated from MD simulations. However, qualitative differences 
  between MD and Markov state dependent model-generated trajectories may limit 
  its usefulness.

@@

  Lyotropic liquid crystals (LLCs) are a class of amphiphilic molecules that
  can be cross-linked into mechanically strong and highly selective nanoporous
  membranes.~\cite{gin_polymerized_2008} They may provide a promising
  alternative to conventional membrane separation techniques by being selective
  based not only on solute size and charge, but on solute chemical
  functionality as well.~\cite{dischinger_application_2017} One can tune the
  functionality of LLC monomers in order to enhance or weaken specific
  solute-membrane interactions.~\cite{dischinger_effect_2017} Two general
  morphologies of LLC membranes are being actively developed. The inverted
  hexagonal, H\textsubscript{II}, LLC phase has densely packed, uniform-sized
  pores on the order of 1 nm in size. A perfectly synthesized
  H\textsubscript{II} phase has an ideal geometry for high throughput
  separations. Unfortunately, aligning the microscopic hexagonal mesophases on
  a large scale as required for high permeability has shown limited
  success.~\cite{feng_scalable_2014,feng_thin_2016} Bicontinuous cubic,
  Q\textsubscript{I}, phase LLC membranes share the uniform and nm-sized chemically
  complex pores of the H\textsubscript{II} phase but its geometry consists of a
  tortuous network of 3D interconnected pores which do not require
  alignment.~\cite{carter_glycerol-based_2012} While its pore structure may
  decrease a Q\textsubscript{I} phase membrane's permeability relative to the
  H\textsubscript{II} phase, it is considerably easier to 
  %MRS10: synthesis implies chemical synthesis, this is really assembly - but produce is more general. 
  %synthesize 
  produce 
  at scale.~\cite{hatakeyama_nanoporous_2010}

@@

  Using a bottom-up modeling approach, we can parameterize single particle
  behavior by extracting 
  %MRS10: their is referring to behavior, which is singular - changing grammar.
  %their 
  particle
  trajectories from MD simulations and studying
  the properties that have the greatest influence on solute dynamics. With this
  information, we can learn how to construct an ensemble of characteristic
  single solute trajectories with these properties which would be useful for
  making long timescale predictions with computational ease and lower
  uncertainties. Fortunately, there is an abundance of information encoded in
  the complex solute trajectories. 
  %MRS: moving clause out from top.
  %Into a model, 
  We can incorporate fluctuations and
  time-dependent correlations in solute position 
  %MRS added:
  into transport models, 
  as well as the length of time
  solutes are trapped. We can add further detail by integrating
  this time series analysis with our knowledge of the primary trapping
  mechanisms as well as the solute's changing chemical environment within the
  heterogeneous membrane.

@@

  We constructed our first model by applying the 
  %MRS10: tweaked
  existing
  rigorous theoretical
  foundation 
  %already developed 
  to describe the motion of 
  particles 
  %MRS10: added specificity
  simulated at the atomistic level
  %exhibiting
  which exhibit
  non-Brownian, or anomalous, transport
  behavior.~\cite{metzler_random_2000,bouchaud_anomalous_1990} The tools
  introduced by fractional calculus are instrumental to this
  theory.~\cite{gorenflo_fractional_1997} They allow us to generalize the
  normally linear diffusion equation to fractional derivative orders, providing
  descriptions of a much more diverse set of behavior, including subdiffusion,
  a type of anomalous diffusion exhibited by solutes in this
  study.~\cite{klages_anomalous_2008}

@@

  We treated our system as an FBM process subordinate to a CTRW, or
  subordinated FBM (sFBM) for short. We fit sFBM models in two ways: First, we
  parameterized solute motion using a single set of parameters fit to a
  hypothesized anomalous diffusion (AD) model. Second, we used a two-state approach
  by extracting two sets of model parameters calculated based on a solute's
  radial distance from the closest pore center. This allows us to include
  different dynamical behavior of solutes 
  %MRS10: added, to show that this is informed by looking at the trajectories.
  previously
  observed within the pore region
  versus within the monomer tails.

  %MRS10: tweaked, also made plural since multiple models are built.
  %Where applicable, the sFBM model 
  Where sufficiently close to observed data, these fitted sFBM models
  provides a way to propagate information
  about solute trajectories gathered at the microsecond time scale to realistic
  experimental time scales. They can also confirm the importance of assumed
  solute-membrane interactions used to define the model's functional form, and
  suggest missing components. They can help researchers formulate hypotheses to
  explain why a certain set of parameters is characteristic to a specific
  solute.

@@

  We evaluate the two modeling approaches by using them to characterize the
  dynamical behavior of the four fastest moving solutes studied in our previous
  work: methanol, urea, ethylene glycol and acetic acid. In addition to moving
  %MRS10: tweaked, may not be quick in absolute terms.
  more 
  quickly
  than other solutes studied previously, 
  allowing them to extensively explore membrane structural space,
  these solutes have a range of chemical functionality and experience each of
  the three trapping mechanisms to different extents.

@@

  \textit{Dwell time distributions:} For subdiffusive transport, the distribution 
  of dwell times is expected to fit a power law distribution 
  proportional to $t^{-1-\alpha}$.~\cite{meroz_toolbox_2015}
  Because we are limited to taking measurements at discrete intervals dictated by the output 
  frequency of our simulation trajectories, we fit the empirical dwell times
  to a discrete power law distribution whose maximum likelihood $\alpha$ 
  parameter we calculated by maximizing the 
  %MRS10: added for precision
  log
  likelihood function: 
  \begin{equation}
	\mathcal{L}(\beta) = -n\ln \zeta(\beta, t_{min}) -
	\beta\sum_{i=1}^{n} \ln t_i 
  \label{eqn:powerlaw_likelihood}
  \end{equation}
  where $\beta = 1 + \alpha$, $t_i$ are collected dwell time data points,
  $n$ the total number of data points, and $\zeta$ is the Hurwitz zeta function
  where $t_{min}$ is the smallest measured value of $t$.~\cite{clauset_power-law_2009}

  In practical applications, the heavy tail of power law distributions can result in 
  arbitrarily long dwell times that are never observed in MD simulations. 
  In order to directly compare our anomalous diffusion model to finite-length MD 
  trajectories we need to bound the dwell time distribution. A standard way of 
  doing this, with easily estimated parameters, is by adding an exponential 
  cut-off to the power law so the dwell time distribution is now proportional 
  to $t^{-1 - \alpha}e^{-\lambda t}$.~\cite{newman_power_2005,clauset_power-law_2009}  % Clauset uses data from other papers and gives a bunch of example where the cut-off is the most favorable model. 
  We determine MLEs of $\alpha$ and $\lambda$ by maximizing the 
  %MRS10:
  log 
  likelihood function:~\cite{clauset_power-law_2009}
  \begin{equation}
    \mathcal{L}(\alpha, \lambda) = n(1 - \alpha)\ln\lambda - n\ln\Gamma(1 - \alpha, t_{min}\lambda) - \alpha\sum_{i=1}^{n}\ln t_i - \lambda\sum_{i=1}^n t_i
  \label{eqn:powerlaw_cutoff_likelihood}
  \end{equation}

@@

  %BJC3: not sure how to cite fbm
  %TODO: cite github page for fbm
  We used the Python package \texttt{fbm} to generate exact simulations of FBM
  %MRS10: add a reference to where your implementation can be found.
  and our own Python implementation of the algorithm by Stoev and Taqqu to
  simulate FLM.~\cite{stoev_simulation_2004} Note that, to our knowledge, there
  are no known exact simulation algorithms for generating FLM trajectories.
  However, the algorithm we used sufficiently approximates draws from the
  marginal L\'evy stable distribution and reasonably approximates the
  correlation structure on MD simulation timescales. We added an empirical
  correction to enhance the accuracy of the correlation structure (see
  Section~\ref{S-section:flm_correlation} of the Supplemental Material for
  validation of the approach).

@@

  A Markov state model (MSM) decomposes a time series into a set of discrete
  states with transitions between states defined by a transition probability
  matrix, $T$.  
  %MRS10: split sentence for readability.
  $T$ describes the conditional probability of moving to a
  specific state given the previously observed
  state.~\cite{pande_everything_2010,wehmeyer_introduction_2018}
%  Software packages such as MSMbuilder~\cite{beauchamp_msmbuilder2:_2011} and 
%  pyEMMA~\cite{scherer_pyemma_2015} provide work flows capable of featurization and
%  dimensional reduction which are often necessary to identify representative states.

@@

  To derive an analytical equation describing the passage time distributions,
  one can frame the problem as a pulse of particles instantaneously released at
  the pore inlet ($z=0$) which moves at a constant velocity, $v$, and spreads
  out as it approaches $L$. This spreading is parameterized by an effective
  diffusivity parameter, $D$. This approach gives results equivalent to if we
  had released each particle individually and then analyzed the positions of
  the ensemble of trajectories as a function of time. The analytical expression
  describing the distribution of first passage times
  is:~\cite{cussler_diffusion:_2009}
  \begin{equation}
  P(t) = -\frac{1}{\sqrt{\pi}}e^{-(L - vt)^2 / (4Dt)}\bigg(-\frac{D(L - vt)}{4(Dt)^{3/2}} - \frac{v}{2\sqrt{Dt}}\bigg)
  \label{eqn:passage_times}
  \end{equation} 
  where the only free parameters for fitting are $v$ and $D$. A derivation of
  Equation~\ref{eqn:passage_times} is given in
  Section~\ref{S-section:fpt_derivation} of the Supplemental Material. We
  calculated the expected value of Equation~\ref{eqn:passage_times} in order to
  get the MFPT.
  %MRS8: check to make sure there is goodness-of-fit checks in the SI for fitting to the analytical model.
  %BJC8: TODO

@@

  \begin{figure}
  \centering
  % /figures/solute_trajectories.py
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{URE_trajectories.pdf}
  \caption{urea}\label{fig:URE_trajectories}
  \end{subfigure}
  % Next 3: /figures/sfbm_distributions
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{GCL_trajectories.pdf}
  \caption{ethylene glycol}\label{fig:GCL_trajectories}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{MET_trajectories.pdf}
  \caption{methanol}\label{fig:MET_trajectories}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{ACH_trajectories.pdf}
  \caption{acetic acid}\label{fig:ACH_trajectories}
  \end{subfigure}
  \caption{
    %MRS10: added
    Three
    representative trajectories generated from each solute exhibit hops
	  between periods of entrapment, characteristic of a CTRW. Solute
	  dynamics show radial dependence, represented by the color at each
	  time point. The longest periods of entrapment typically occur when
	  solutes drift far from the pore center and into the tails.
  }\label{fig:solute_trajectories}
  \end{figure}

  \begin{figure}
  \centering
  % Next 3: /figures/sfbm_distributions
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison_anomalous_URE.pdf}
  \caption{}\label{fig:hop_distribution_comparison}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{URE_powerlaw.pdf}
  \caption{}\label{fig:powerlaw}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{URE_hop_acf.pdf}
  \caption{}\label{fig:hop_acf}
  \end{subfigure}
  \caption{Using urea as an example, (a) the distribution of hop lengths can be
	  fit by a Gaussian 
          ``as well as'' is ambiguous, can mean ``in addition to'' or ``similar in quality to''
          %as well as 
          almost as well as
          a more general L\'evy stable
	  distribution, though the L\'evy stable distribution does a better job of
	  capturing the heavier tails and increased density near 0. We explore
	  models fit to both distributions, as Gaussian hops are more convenient
	  to model. (b) The distribution of dwell times is fit well by a power
	  law but it over-estimates the probability density at long dwell
	  times. A power law truncated with an exponential cut-off better
	  describes the probability of long dwell times in our simulations. (c)
	  The hops are negatively correlated to their previous hop. In
	  combination, (a) -- (c) support modeling solutes as either
	  subordinated fractional Brownian or L\'evy motion. All other solutes
	  show similar distributions and autocorrelation functions (see
	  Figure~\ref{S-fig:anticorrelated_hops} of the Supplemental Material).
  }\label{fig:anticorrelated_hops}
  \end{figure}

@@

  We observe that in some cases, solute trajectories extracted from our MD
  simulations display non-stationary behavior. We defined the perceived
  equilibration time point for each solute based on the time at which the
  number of solutes inside the pores and tails stabilized
  (Figure~\ref{S-fig:equilibration}).  With this definition, we observe
  evidence of non-stationary solute behavior after the perceived equilibration
  point, on the $\mu$s timescale. In Table~\ref{table:stationarity}, we compare
  the MSD of the solutes calculated using trajectory data from the first and
  second halves of the ``equilibrated" simulation time. Ethylene glycol and
  methanol show considerable differences between the MSDs of the two halves.
  Urea and acetic acid appear to demonstrate satisfactory stationarity. The
  shapes of 
  %MRS10: which of the solutes does ``their'' refer to?  Be more explicit.
  their MSD curves are also very similar (see
  Figure~\ref{S-fig:msd_comparison} of the Supplemental Material).

@@

  We obtain reasonable predictions of the MD simulated MSDs when we
  parameterize AD models with all available data after the perceived
  equilibration time. The MSD curves generated from 
  %MRS10: added
  both 
  %MRS10: more grammatically correct to say ``one and two'' instead of ``1 and 2''
  the one and two mode models
  are overlayed with the MD simulated MSDs for comparison in
  Figure~\ref{fig:anomalous_msds}. The associated parameters for the one and two
  mode models are presented in Figures~\ref{fig:1mode_parameters}
  and~\ref{fig:2mode_parameters}.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{1mode_AD_hops.pdf}
  \caption{}\label{fig:1mode_AD_hops}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{1mode_AD_dwells.pdf}
  \caption{}\label{fig:1mode_AD_dwells}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{hurst_barchart.pdf}
  \caption{}\label{fig:hurst_barchart}
  \end{subfigure}
  \caption{The parameters of the one mode model reveal differences in dynamics
	  between solutes. (a) We parameterized Gaussian,
	  $\mathcal{N}(\sigma)$, and L\'evy stable, $L(\sigma, \alpha_h)$,
	  distributions to describe solute hop lengths. We assume the mean
	  ($\mu$) to be zero for these distributions and no
	  skewness ($\beta = 0$) in the L\'evy stable distributions. High
	  values of $\sigma$ and lower values of $\alpha_h$ result in larger
	  hops. (b) We parameterized a pure power law, $P(\alpha)$, and a
	  truncated power law, $P_T(\alpha, \lambda)$, distribution to describe
	  solute dwell times. Lower values of $\alpha$ lead to heavier power
	  law tails and higher values of $\lambda$ truncate the distribution at
	  lower dwell times. (c) Finally, we parameterized the hop
	  autocorrelation function, $\gamma(H)$, to describe the degree of
	  correlation between hops. 
          %MRS10: tweaked
          Simulations with higher values of $H$ display behavior closer to
	  the Brownian limit.}\label{fig:1mode_parameters}
  \end{figure}

@@

  The deviation between the one mode MSD predictions and MD are primarily due
  to differences in their curvature at long time lags. The shape and magnitude
  of the predicted curves appears accurate relative to MD at short time lags.
  However, the modeled trajectories undershoot the mean MD MSD as the time lag
  increases. As discussed in the previous section, long time positional
  anti-correlation, on the order of hundreds of ns, may not exist in the MD
  system. Eventual loss of correlation would result in a shift from sub-linear
  to linear MSD behavior, as observed in the MD trajectories. Acetic acid
  exemplifies this point. At first glance, acetic acid's predicted MSD appears
  to match the curvature of MD quite well, but closer examination reveals that
  the MD MSD curve 
  %MRS10: being a bit more cautious about claims.
  %actually shifts 
  may actually shift
  from a sub-linear to a linear regime around
  500 ns.

@@

  %MRS10: tweak
  %Turning first to 
  Examining first
  the parameters of the one mode model, we can begin to break
  down the trends in solute MSDs. The parameters belonging to ethylene glycol
  and methanol are relatively similar, which is consistent with their similar MD
  MSDs
  %MRS10: added
  and qualitatively similar MD timeseries.
  Relative to ethylene glycol, methanol tends to stay trapped for less
  time and takes larger hops but the most substantial difference is with
  respect to their Hurst parameters. Methanol has the lowest $H$ of all the
  solutes studied because it spends the majority of its time outside the pore
  region where collisions with tails are frequent.
% BJC: made more to the point
%  In fact, methanol has the lowest
%  $H$ of all the solutes studied. It is possible that we have underestimated $H$ since 
%  it appears that our model under-predicts the methanol MSD curve relative
%  to other solutes. However, it remains plausible that methanol does have a low $H$ 
%  value because it spends the majority of its time outside the pore region where collisions
%  with tails are frequent. 
  Urea has the third highest MSD which is primarily a consequence of more
  frequent and longer dwell times (lower $\alpha$ and $\lambda$). Urea's hop
  lengths ($\sigma$) and correlation ($H$) are comparable to ethylene glycol
  and methanol. Acetic acid has the smallest MSD among the solutes studied due
  to longer periods of entrapment and shorter hops. Its trapping behavior is
  parameterized by an $\alpha$ value significantly lower than other solutes,
  but an intermediate $\lambda$ value, suggesting it experiences many
  medium-length periods of entrapment. Its hops are smaller but are slightly
  compensated by a heavier tailed distribution (lower $\alpha_h$) than the
  other solutes.

@@

  To create an MSDDM for each solute, we determined the state sequence associated
  with each solute trajectory 
  %MRS10: added for clarity to make sure people who how the state sequence was determined.
  based on the geometric indicators of the states indicated in Table~\ref{table:states}.
  and then generated emission distributions of fluctuations
  within each state as well as transitions between states. In theory, one could 
  parameterize separate transition distributions for those which occur in the tails 
  versus in the pores, however this would lead to a broken correlation structure 
  similar to that seen in the two mode AD models.

>>>>

  Next, we generalize Markov state models, treating the
  configurational states of the system as a Markov process where each
  state has distinct transport properties. For each state and transition
  between states, we parameterize the distribution and temporal correlation 
  structure of positional fluctuations as a means of characterization and to 
  allow us to predict solute MSDs. We show that both models reasonably reproduce
  the MSDs calculated from MD simulations. However, qualitative differences 
  between MD and Markov state dependent model-generated trajectories may limit 
  its usefulness.

@@

  Lyotropic liquid crystals (LLCs) are a class of amphiphilic molecules that
  can be cross-linked into mechanically strong and highly selective nanoporous
  membranes.~\cite{gin_polymerized_2008} They may provide a promising
  alternative to conventional membrane separation techniques by being selective
  based not only on solute size and charge, but on solute chemical
  functionality as well.~\cite{dischinger_application_2017} One can tune the
  functionality of LLC monomers in order to enhance or weaken specific
  solute-membrane interactions.~\cite{dischinger_effect_2017} Two general
  morphologies of LLC membranes are being actively developed. The inverted
  hexagonal, H\textsubscript{II}, LLC phase has densely packed, uniform-sized
  pores on the order of 1 nm in size. A perfectly synthesized
  H\textsubscript{II} phase has an ideal geometry for high throughput
  separations. Unfortunately, aligning the microscopic hexagonal mesophases on
  a large scale as required for high permeability has shown limited
  success.~\cite{feng_scalable_2014,feng_thin_2016} Bicontinuous cubic,
  Q\textsubscript{I}, phase LLC membranes share the uniform and nm-sized chemically
  complex pores of the H\textsubscript{II} phase but its geometry consists of a
  tortuous network of 3D interconnected pores which do not require
  alignment.~\cite{carter_glycerol-based_2012} While its pore structure may
  decrease a Q\textsubscript{I} phase membrane's permeability relative to the
  H\textsubscript{II} phase, it is considerably easier to produce 
  at scale.~\cite{hatakeyama_nanoporous_2010}

@@

  Using a bottom-up modeling approach, we can parameterize single particle
  behavior by extracting particle trajectories from MD simulations and studying
  the properties that have the greatest influence on solute dynamics. With this
  information, we can learn how to construct an ensemble of characteristic
  single solute trajectories with these properties which would be useful for
  making long timescale predictions with computational ease and lower
  uncertainties. Fortunately, there is an abundance of information encoded in
  the complex solute trajectories. We can incorporate fluctuations and 
  time-dependent correlations in solute position into transport models, 
  as well as the length of time solutes are trapped. We can add further detail by integrating
  this time series analysis with our knowledge of the primary trapping
  mechanisms as well as the solute's changing chemical environment within the
  heterogeneous membrane.

@@

  We constructed our first model by applying the existing rigorous theoretical
  foundation which describes the motion of particles that exhibit non-Brownian,
  or anomalous, transport behavior.~\cite{metzler_random_2000,bouchaud_anomalous_1990} The tools
  introduced by fractional calculus are instrumental to this
  theory.~\cite{gorenflo_fractional_1997} They allow us to generalize the
  normally linear diffusion equation to fractional derivative orders, providing
  descriptions of a much more diverse set of behavior, including subdiffusion,
  a type of anomalous diffusion exhibited by solutes in this
  study.~\cite{klages_anomalous_2008}

@@

  We treated our system as an FBM process subordinate to a CTRW, or
  subordinated FBM (sFBM) for short. We fit sFBM models in two ways: First, we
  parameterized solute motion using a single set of parameters fit to a
  hypothesized anomalous diffusion (AD) model. Second, we used a two-state approach
  by extracting two sets of model parameters calculated based on a solute's
  radial distance from the closest pore center. This allows us to include
  the different dynamical behavior of solutes previously observed within the
  pore region versus within the monomer tails.

  Where sufficiently close to observed data, these fitted sFBM models
  provide a way to propagate information about solute trajectories gathered 
  at the microsecond time scale to realistic experimental time scales. 
  They can also confirm the importance of assumed solute-membrane interactions 
  used to define the model's functional form, and
  suggest missing components. They can help researchers formulate hypotheses to
  explain why a certain set of parameters is characteristic to a specific
  solute.

@@

  We evaluate the two modeling approaches by using them to characterize the
  dynamical behavior of the four fastest moving solutes studied in our previous
  work: methanol, urea, ethylene glycol and acetic acid. In addition to moving
  more quickly than other solutes studied previously, allowing them to extensively
  explore membrane structural space, these solutes have a range of chemical 
  functionality and experience each of the three trapping mechanisms to different extents.

@@

  \textit{Dwell time distributions:} For subdiffusive transport, the distribution 
  of dwell times is expected to fit a power law distribution 
  proportional to $t^{-1-\alpha}$.~\cite{meroz_toolbox_2015}
  Because we are limited to taking measurements at discrete intervals dictated by the output 
  frequency of our simulation trajectories, we fit the empirical dwell times
  to a discrete power law distribution whose maximum likelihood $\alpha$ 
  parameter we calculated by maximizing the log likelihood function: 
  \begin{equation}
	\mathcal{L}(\beta) = -n\ln \zeta(\beta, t_{min}) -
	\beta\sum_{i=1}^{n} \ln t_i 
  \label{eqn:powerlaw_likelihood}
  \end{equation}
  where $\beta = 1 + \alpha$, $t_i$ are collected dwell time data points,
  $n$ the total number of data points, and $\zeta$ is the Hurwitz zeta function
  where $t_{min}$ is the smallest measured value of $t$.~\cite{clauset_power-law_2009}

  In practical applications, the heavy tail of power law distributions can result in 
  arbitrarily long dwell times that are never observed in MD simulations. 
  In order to directly compare our anomalous diffusion model to finite-length MD 
  trajectories we need to bound the dwell time distribution. A standard way of 
  doing this, with easily estimated parameters, is by adding an exponential 
  cut-off to the power law so the dwell time distribution is now proportional 
  to $t^{-1 - \alpha}e^{-\lambda t}$.~\cite{newman_power_2005,clauset_power-law_2009}  % Clauset uses data from other papers and gives a bunch of example where the cut-off is the most favorable model. 
  We determine MLEs of $\alpha$ and $\lambda$ by maximizing the log 
  likelihood function:~\cite{clauset_power-law_2009}
  \begin{equation}
    \mathcal{L}(\alpha, \lambda) = n(1 - \alpha)\ln\lambda - n\ln\Gamma(1 - \alpha, t_{min}\lambda) - \alpha\sum_{i=1}^{n}\ln t_i - \lambda\sum_{i=1}^n t_i
  \label{eqn:powerlaw_cutoff_likelihood}
  \end{equation}

@@

  We used the Python package \texttt{fbm}~\cite{flynn_fbm_2019} to generate exact simulations of FBM
  and our own Python implementation (see Table~\ref{S-table:python_scripts} of
  the Supplemental Material) of the algorithm by Stoev and Taqqu to
  simulate FLM.~\cite{stoev_simulation_2004} Note that, to our knowledge, there
  are no known exact simulation algorithms for generating FLM trajectories.
  However, the algorithm we used sufficiently approximates draws from the
  marginal L\'evy stable distribution and reasonably approximates the
  correlation structure on MD simulation timescales. We added an empirical
  correction to enhance the accuracy of the correlation structure (see
  Section~\ref{S-section:flm_correlation} of the Supplemental Material for
  validation of the approach).

@@

  A Markov state model (MSM) decomposes a time series into a set of discrete
  states with transitions between states defined by a transition probability
  matrix, $T$.  $T$ describes the conditional probability of moving to a
  specific state given the previously observed
  state.~\cite{pande_everything_2010,wehmeyer_introduction_2018}
%  Software packages such as MSMbuilder~\cite{beauchamp_msmbuilder2:_2011} and 
%  pyEMMA~\cite{scherer_pyemma_2015} provide work flows capable of featurization and
%  dimensional reduction which are often necessary to identify representative states.

@@

  To derive an analytical equation describing the passage time distributions,
  one can frame the problem as a pulse of particles instantaneously released at
  the pore inlet ($z=0$) which moves at a constant velocity, $v$, and spreads
  out as it approaches $L$. This spreading is parameterized by an effective
  diffusivity parameter, $D$. This approach gives results equivalent to if we
  had released each particle individually and then analyzed the positions of
  the ensemble of trajectories as a function of time. The analytical expression
  describing the distribution of first passage times
  is:~\cite{cussler_diffusion:_2009}
  \begin{equation}
  P(t) = -\frac{1}{\sqrt{\pi}}e^{-(L - vt)^2 / (4Dt)}\bigg(-\frac{D(L - vt)}{4(Dt)^{3/2}} - \frac{v}{2\sqrt{Dt}}\bigg)
  \label{eqn:passage_times}
  \end{equation} 
  where the only free parameters for fitting are $v$ and $D$. A derivation of
  Equation~\ref{eqn:passage_times} is given in
  Section~\ref{S-section:fpt_derivation} of the Supplemental Material. We
  calculated the expected value of Equation~\ref{eqn:passage_times} in order to
  get the MFPT.

@@

  \begin{figure}
  \centering
  % /figures/solute_trajectories.py
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{URE_trajectories.pdf}
  \caption{urea}\label{fig:URE_trajectories}
  \end{subfigure}
  % Next 3: /figures/sfbm_distributions
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{GCL_trajectories.pdf}
  \caption{ethylene glycol}\label{fig:GCL_trajectories}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{MET_trajectories.pdf}
  \caption{methanol}\label{fig:MET_trajectories}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{ACH_trajectories.pdf}
  \caption{acetic acid}\label{fig:ACH_trajectories}
  \end{subfigure}
  \caption{Three representative trajectories generated from each solute exhibit hops
	  between periods of entrapment, characteristic of a CTRW. Solute
	  dynamics show radial dependence, represented by the color at each
	  time point. The longest periods of entrapment typically occur when
	  solutes drift far from the pore center and into the tails.
  }\label{fig:solute_trajectories}
  \end{figure}

  \begin{figure}
  \centering
  % Next 3: /figures/sfbm_distributions
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison_anomalous_URE.pdf}
  \caption{}\label{fig:hop_distribution_comparison}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{URE_powerlaw.pdf}
  \caption{}\label{fig:powerlaw}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{URE_hop_acf.pdf}
  \caption{}\label{fig:hop_acf}
  \end{subfigure}
  \caption{Using urea as an example, (a) the distribution of hop lengths can be
	  fit by a Gaussian in addition to a more general L\'evy stable
	  distribution, though the L\'evy stable distribution does a better job of
	  capturing the heavier tails and increased density near 0. We explore
	  models fit to both distributions, as Gaussian hops are more convenient
	  to model. (b) The distribution of dwell times is fit well by a power
	  law but it over-estimates the probability density at long dwell
	  times. A power law truncated with an exponential cut-off better
	  describes the probability of long dwell times in our simulations. (c)
	  The hops are negatively correlated to their previous hop. In
	  combination, (a) -- (c) support modeling solutes as either
	  subordinated fractional Brownian or L\'evy motion. All other solutes
	  show similar distributions and autocorrelation functions (see
	  Figure~\ref{S-fig:anticorrelated_hops} of the Supplemental Material).
  }\label{fig:anticorrelated_hops}
  \end{figure}

@@

  We observe that in some cases, solute trajectories extracted from our MD
  simulations display non-stationary behavior. We defined the perceived
  equilibration time point for each solute based on the time at which the
  number of solutes inside the pores and tails stabilized
  (Figure~\ref{S-fig:equilibration}).  With this definition, we observe
  evidence of non-stationary solute behavior after the perceived equilibration
  point, on the $\mu$s timescale. In Table~\ref{table:stationarity}, we compare
  the MSD of the solutes calculated using trajectory data from the first and
  second halves of the ``equilibrated" simulation time. Ethylene glycol and
  methanol show considerable differences between the MSDs of the two halves.
  Urea and acetic acid appear to demonstrate satisfactory stationarity. The
  shapes of urea and acetic acid's MSD curves are also very similar (see
  Figure~\ref{S-fig:msd_comparison} of the Supplemental Material).

@@

  We obtain reasonable predictions of the MD simulated MSDs when we
  parameterize AD models with all available data after the perceived
  equilibration time. The MSD curves generated from both the one and 
  two mode models are overlayed with the MD simulated MSDs for comparison in
  Figure~\ref{fig:anomalous_msds}. The associated parameters for the one and two
  mode models are presented in Figures~\ref{fig:1mode_parameters}
  and~\ref{fig:2mode_parameters}.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{1mode_AD_hops.pdf}
  \caption{}\label{fig:1mode_AD_hops}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{1mode_AD_dwells.pdf}
  \caption{}\label{fig:1mode_AD_dwells}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{hurst_barchart.pdf}
  \caption{}\label{fig:hurst_barchart}
  \end{subfigure}
  \caption{The parameters of the one mode model reveal differences in dynamics
	  between solutes. (a) We parameterized Gaussian,
	  $\mathcal{N}(\sigma)$, and L\'evy stable, $L(\sigma, \alpha_h)$,
	  distributions to describe solute hop lengths. We assume the mean
	  ($\mu$) to be zero for these distributions and no
	  skewness ($\beta = 0$) in the L\'evy stable distributions. High
	  values of $\sigma$ and lower values of $\alpha_h$ result in larger
	  hops. (b) We parameterized a pure power law, $P(\alpha)$, and a
	  truncated power law, $P_T(\alpha, \lambda)$, distribution to describe
	  solute dwell times. Lower values of $\alpha$ lead to heavier power
	  law tails and higher values of $\lambda$ truncate the distribution at
	  lower dwell times. (c) Finally, we parameterized the hop
	  autocorrelation function, $\gamma(H)$, to describe the degree of
	  correlation between hops. Simulations with higher values of $H$ display
	  behavior closer to the Brownian limit.}\label{fig:1mode_parameters}
  \end{figure}

@@

  The deviation between the one mode MSD predictions and MD are primarily due
  to differences in their curvature at long time lags. The shape and magnitude
  of the predicted curves appears accurate relative to MD at short time lags.
  However, the modeled trajectories undershoot the mean MD MSD as the time lag
  increases. As discussed in the previous section, long time positional
  anti-correlation, on the order of hundreds of ns, may not exist in the MD
  system. Eventual loss of correlation would result in a shift from sub-linear
  to linear MSD behavior, as observed in the MD trajectories. Acetic acid
  exemplifies this point. At first glance, acetic acid's predicted MSD appears
  to match the curvature of MD quite well, but closer examination reveals that
  the MD MSD curve may actually shift from a sub-linear to a linear regime around
  500 ns.

@@

  Examining first the parameters of the one mode model, we can begin to break
  down the trends in solute MSDs. The parameters belonging to ethylene glycol
  and methanol are relatively similar, which is consistent with their similar MD
  MSDs and qualitatively similar MD timeseries.
  Relative to ethylene glycol, methanol tends to stay trapped for less
  time and takes larger hops but the most substantial difference is with
  respect to their Hurst parameters. Methanol has the lowest $H$ of all the
  solutes studied because it spends the majority of its time outside the pore
  region where collisions with tails are frequent.
  Urea has the third highest MSD which is primarily a consequence of more
  frequent and longer dwell times (lower $\alpha$ and $\lambda$). Urea's hop
  lengths ($\sigma$) and correlation ($H$) are comparable to ethylene glycol
  and methanol. Acetic acid has the smallest MSD among the solutes studied due
  to longer periods of entrapment and shorter hops. Its trapping behavior is
  parameterized by an $\alpha$ value significantly lower than other solutes,
  but an intermediate $\lambda$ value, suggesting it experiences many
  medium-length periods of entrapment. Its hops are smaller but are slightly
  compensated by a heavier tailed distribution (lower $\alpha_h$) than the
  other solutes.

@@

  To create an MSDDM for each solute, we determined the state sequence associated
  with each solute trajectory based on the geometric indicators of the states 
  indicated in Table~\ref{table:states}. We then generated emission distributions of fluctuations
  within each state as well as transitions between states. In theory, one could 
  parameterize separate transition distributions for those which occur in the tails 
  versus in the pores, however this would lead to a broken correlation structure 
  similar to that seen in the two mode AD models.
